---
title: "Arizona Data Comparison"
author: "Joe Champion, Thea Sukianto"
date: "August 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Helper Functions/Libraries
```{r message=FALSE}
source('R/data_analysis.R')
library(dplyr)
library(ggplot2)
```

### Load Correct (manually cleaned) Arizona Data

* Renames 'collar', 'date', 'distancetr', and 'RATE' columns to 'GPS', 'Date', 'Distance', and 'Rate' respectively so names match with candidate data 

```{r}
df_correct <- read.csv('inst/extdata/correct.csv', stringsAsFactors = F) %>% 
  dplyr::rename(GPS = collar) %>% 
  dplyr::rename(Date = date) %>% 
  dplyr::rename(Distance = distancetr) %>% 
  dplyr::rename(Rate = RATE) %>% 
  dplyr::mutate(Date = as.Date(Date))
```

### Load Candidate (cleaned by animaltracker) Arizona Data

```{r}
df_candidate <- read.csv('inst/extdata/candidate.csv', stringsAsFactors = F) %>% 
  dplyr::mutate(Date = as.Date(Date))
```

## Calculate Summary Statistics

```{r}
dfs <- compare_summarise_data(df_correct, 
                       df_candidate, 
                       'inst/extdata/processed/gps_arizona.csv', 
                       'inst/extdata/processed/date_arizona.csv'
                       )
```

### Summary by GPS Collar

* **Close match** between latitude, longitude, course
* Candidate has **less rows**
* **Underestimates** distance, rate **(less variable)**
* **Overestimates** elevation **(more variable)**

```{r}
summary(dfs$GPS)
```

#### Number of Observations

* Candidate has approximately 90-170 **less rows** (out of ~6400 - 7500 per GPS collar)

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "n", "inst/extdata/processed/n_GPS_arizona.png")
```

#### Mean Latitude

* Differences are on the order of **~1x10^-5** on average

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "meanLat", "inst/extdata/processed/meanLat_GPS_arizona.png")
```

#### Mean Longitude

* Differences are on the order of **~1x10^-5 - 1x10^-4** on average

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "meanLong", "inst/extdata/processed/meanLong_GPS_arizona.png")
```

#### Mean Distance

* Candidate **underestimates** distance on average by ~4.5 - 7

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "meanDist", "inst/extdata/processed/meanDist_GPS_arizona.png")
```

#### Mean Course

* Differences are on the order of **~1x10^2 - 1x10^1** on average

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "meanCourse", "inst/extdata/processed/meanCourse_GPS_arizona.png")
```

#### Mean Rate

* Candidate **underestimates** rate on average by ~0.20 - 0.28

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "meanRate", "inst/extdata/processed/meanRate_GPS_arizona.png")
```

#### Mean Elevation 

* Candidate **overestimates** elevation on average by ~19 - 26

```{r message=FALSE}
violin_compare(dfs$GPS, GPS, "meanElev", "inst/extdata/processed/meanElev_GPS_arizona.png")
```

### Summary by Date

* **Close match** between latitude, longitude, course
* Candidate has **less dates**
* **Underestimates** distance, rate **(less variable)**
* **Overestimates** elevation **(more variable)**

```{r}
summary(dfs$Date)
```

#### Number of Observations

* Candidate tends to have **less rows** per date (difference between correct and candidate: ~ -1 - 60)

```{r message=FALSE}
violin_compare(dfs$Date, Date, "n", "inst/extdata/processed/n_Date_arizona.png")
```

#### Mean Latitude

* Differences are on the order of **~1x10-4** on average

```{r message=FALSE}
violin_compare(dfs$Date, Date, "meanLat", "inst/extdata/processed/meanLat_Date_arizona.png")
```

#### Mean Longitude

* Differences are on the order of **~1x10-4** on average

```{r message=FALSE}
violin_compare(dfs$Date, Date, "meanLong", "inst/extdata/processed/meanLong_Date_arizona.png")
```

#### Mean Distance

* Candidate **underestimates** distance on average by ~3.3 - 8.6

```{r message=FALSE}
violin_compare(dfs$Date, Date, "meanDist", "inst/extdata/processed/meanDist_Date_arizona.png")
```

#### Mean Course

* Differences are **~ +/- 1** on average**

```{r message=FALSE}
violin_compare(dfs$Date, Date, "meanCourse", "inst/extdata/processed/meanCourse_Date_arizona.png")
```

#### Mean Rate

* Candidate **underestimates** rate on average by ~0.07 - 0.38

```{r message=FALSE}
violin_compare(dfs$Date, Date, "meanRate", "inst/extdata/processed/meanRate_Date_arizona.png")
```

#### Mean Elevation

* Candidate **overestimates** elevation on average by ~34 - 10

```{r message=FALSE}
violin_compare(dfs$Date, Date, "meanElev", "inst/extdata/processed/meanElev_Date_arizona.png")
```

### Summary by GPS and Date

* **Close match** between latitude, longitude
* **Underestimates and overestimates** number of rows and course, but difference is a **close match on average**
* **Tends to underestimate** distance, rate **(less variable)**
* **Overestimates** elevation (variability fluctuates)

```{r message=FALSE}
daily_summary <- compare_summarise_daily(df_correct, df_candidate, "inst/extdata/processed/gps_daily_arizona.csv")

summary(daily_summary)
```

#### Number of Observations

* Differs by **~ -5 - 8** rows (correct - candidate) per GPS-date pair

```{r message=FALSE}
  
  correct_n <- df_correct %>% 
    dplyr::group_by(GPS, Date) %>% 
    dplyr::summarise(n = n()) %>% 
    dplyr::mutate(Data = "Correct")
  
  candidate_n <- df_candidate %>% 
    dplyr::group_by(GPS, Date) %>% 
    dplyr::summarise(n = n()) %>% 
    dplyr::mutate(Data = "Candidate")
  
  plot_data <- dplyr::bind_rows(correct_n, candidate_n)
  
  ggplot(plot_data, aes(x = Date, y = n, group = Data, color = Data)) +
    geom_line() +
    scale_color_discrete(guide = guide_legend(reverse = T)) +
    facet_wrap(vars(GPS))
  

```

#### Mean Latitude

* Differences are on the order of **~1x10-4** on average

```{r message=FALSE}
line_compare(df_correct, df_candidate, Latitude, "inst/extdata/processed/meanLat_line_arizona.png")
```

#### Mean Longitude

* Differences are on the order of **~1x10-4** on average

```{r message=FALSE}
line_compare(df_correct, df_candidate, Longitude, "inst/extdata/processed/meanLong_line_arizona.png")
```

#### Mean Distance

* Differs by **~ -4 - 13.6** on average
* Possible influential points:

```{r message=FALSE}

knitr::kable(daily_summary %>% 
               dplyr::filter((meanDistDiff > quantile(meanDistDiff, 0.75) + 1.5*IQR(meanDistDiff)) |
                             (meanDistDiff < quantile(meanDistDiff, 0.25) - 1.5*IQR(meanDistDiff))) %>%
               dplyr::select(GPS, Date, meanDistDiff))

line_compare(df_correct, df_candidate, Distance, "inst/extdata/processed/meanDist_line_arizona.png")
```

#### Mean Course

* Differences are **~ +/- 6** on average
* Possible influential points:

```{r message=FALSE}
knitr::kable(daily_summary %>% 
               dplyr::filter((meanCourseDiff > quantile(meanCourseDiff, 0.75) + 1.5*IQR(meanCourseDiff)) |
                             (meanCourseDiff < quantile(meanCourseDiff, 0.25) - 1.5*IQR(meanCourseDiff))) %>%
               dplyr::select(GPS, Date, meanCourseDiff))

line_compare(df_correct, df_candidate, Course, "inst/extdata/processed/meanCourse_line_arizona.png")
```

#### Mean Rate

* Differs by **~ -0.16 - 0.64** on average
* Possible influential points:

```{r message=FALSE}
knitr::kable(daily_summary %>% 
               dplyr::filter((meanRateDiff > quantile(meanRateDiff, 0.75) + 1.5*IQR(meanRateDiff)) |
                             (meanRateDiff < quantile(meanRateDiff, 0.25) - 1.5*IQR(meanRateDiff))) %>%
               dplyr::select(GPS, Date, meanRateDiff))

line_compare(df_correct, df_candidate, Rate, "inst/extdata/processed/meanRate_line_arizona.png")
```

#### Mean Elevation

* Candidate **overestimates** elevation on average by ~3 - 43
* Possible influential points:

```{r message=FALSE}
knitr::kable(daily_summary %>% 
               dplyr::filter((meanElevDiff > quantile(meanElevDiff, 0.75) + 1.5*IQR(meanElevDiff)) |
                             (meanElevDiff < quantile(meanElevDiff, 0.25) - 1.5*IQR(meanElevDiff))) %>%
               dplyr::select(GPS, Date, meanElevDiff))

line_compare(df_correct, df_candidate, Elevation, "inst/extdata/processed/meanElev_line_arizona.png")
```
